Map{R}educe in {MPI} for Large-scale Graph Algorithms

### 2010-05-28 paper/results.tex
More.
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@625 3fc257bb-8550-0410-81b3-a5c103da83f7

In this section, we present performance results for the MapReduce
graph algorithms of the preceeding section, implemented as small C++
programs calling our MR-MPI library.
The benchmarks were run on a medium-sized Linux cluster made up of 2 GHz
dual-core AMD Opteron processors with a Myrinet network.  Most importantly
for MR-MPI, each node of the cluster has one local disk, which is used
for out-of-core operations in MR-MPI.  To avoid contention for disk I/O, we 
ran all experiments with one MPI process per node.

@@

\begin{table}
\begin{tabular}{|l|c|c|c|c|c|c|c|}
\hline
Data & \# of    & \# of & Maximum \\
Set  & vertices & edges & vertex degree\\
\hline
RMAT-20 (small)   &$2^{20} \approx 1M$ & $2^{23} \approx 8M$ &   \\
RMAT-24 (medium)  &$2^{24} \approx 17M$ & $2^{27} \approx 134M$ &   \\
RMAT-28 (large)   &$2^{28} \approx 268M$ & $2^{31} \approx 2B$&   \\
%RMAT-32 (x-large) &$2^{32} \approx 4B$ & $2^{35} \approx 34B$ &   \\
\hline
\end{tabular}
\caption{Characteristics of R-MAT input data for graph algorithm
experiments.}
\label{t:rmats}
\end{table}

The resulting timings give a sense of the inherent scalability of the
MapReduce algorithms as graph size grows on a fixed number of
processors, and of the parallel scalability for computing on a graph
of fixed size on a growing number of processors.  Where available, we
compare the MapReduce algorithm with other parallel implementations.

@@

\begin{figure}[h!]
\includegraphics[width=\textwidth]{fig_rmat.pdf}
\caption{Performance of R-MAT generation algorithm (Figure~\ref{fig:rmat2}).}
\label{f:rmat}
\end{figure}

@@

\begin{figure}[h!]
\includegraphics[width=\textwidth]{fig_pagerank.pdf}
\caption{Comparison of PageRank implementations using MapReduce,
Trilinos, and MTGL on R-MAT data sets.}
\label{f:pr}
\end{figure}

\subsection{Triangle finding results}
Discuss triangle finding times.

@@

\subsection{Connected Components}
Discuss Luby maximally independent sets times.

@@

The execution times for the maximally independent set algorithm
(Figure~\ref{fig:luby}) are shown in Figure~\ref{f:luby}.  Like the R-MAT
generation results, superlinear speed-up of the algorithm occurs for RMAT-20
and RMAT-24, as more of the graph fits into processor memory and less file
I/O is needed.  For RMAT-28, the algorithm requires significant out-of-core
operations. In this case, parallel efficiency is nearly perfect going from 
8 to 32 processors; some loss of efficiency is seen at 64 processors,
where the parallel efficiency compared to the 8-processor case is ??\%.

@@

\begin{figure}[h!]
\includegraphics[width=\textwidth]{fig_luby.pdf}
\caption{Performance of maximally independent set algorithm (Figure~\ref{fig:luby}).}
\label{f:luby}
\end{figure}

@@

\begin{figure}[h!]
\includegraphics[width=\textwidth]{fig_sssp.pdf}
\caption{Execution times for SSSP using MR-MPI with R-MAT matrices.  
Both the original algorithm (Figure~\ref{fig:sssp}) and the enhanced 
algorithm (Figure~\ref{fig:sssp2}) are included.}
\label{f:sssp}
\end{figure}

@@

\begin{figure}[h!]
\includegraphics[width=\textwidth]{fig_ssspA.pdf}
\caption{SSSP using MR-MPI for WebGraphA with
13.2M vertices and 31.9M edges.  Runtimes using Hadoop and PBGL
are also shown.}
\label{f:ssspA}
\end{figure}

\begin{table}
\begin{tabular}{|l|c|r|}
\hline
Implementation & Number of Processes & SSSP Execution Time \\
\hline
Hadoop & 48  & 38,925 secs.\\
MR-MPI original (Figure~\ref{fig:sssp}) & 48 &  13,505 secs.\\
MR-MPI enhanced (Figure~\ref{fig:sssp2}) & 48 &  8,031 secs.\\
XMT/C  & 32 &  37 secs.\\
%Out-of-core MR-MPI & 96 &   8,358 secs.\\
%Out-of-core MR-MPI & 64 &  12,882 secs.\\
%Out-of-core MR-MPI & 100 &  6,280 secs.\\
\hline
\end{tabular}
\caption{Execution times for SSSP with {WebGraphB}.}
\label{t:ssspB}
\end{table}

\subsection{Scalability to large numbers of processors}
Finally, we demonstrate the scalability of MR-MPI to large numbers of 
processors.  
The MR-MPI implementation was run on the Redstorm and Thunderbird
parallel computers
Because these systems do not have local disks for each processor, we selected 
a data set and page sizes that fit in their memory, so out-of-core operations
were not used.  For these experiments, we used an R-MAT data set with 
with $2^{25}$ vertices and $2^{28}$ edges, with parameters given in
Table~\ref{t:rmat}.  We ran both PageRank and Connected Components
algorithms.

\begin{table}
\begin{tabular}{|l|c|c|c|c|c|c|c|}
\hline
Data & R-MAT  & R-MAT  & R-MAT  & R-MAT  & \# of    & \# of & Maximum \\
Set  & a      & b      & c      & d      & vertices & edges & vertex degree\\
\hline
nice  & 0.45 & 0.15 & 0.15 & 0.25 & $2^{25}$ & $2^{28}$ & 1108 \\
nasty & 0.57 & 0.19 & 0.19 & 0.05 & $2^{25}$ & $2^{28}$ & 230,207\\
\hline
\end{tabular}
\caption{Characteristics of R-MAT input data for PageRank and Connected
Components scalability experiments.}
\label{t:rmat}
\end{table}

@@

\begin{figure}[h!]
\includegraphics[width=\textwidth]{fig_pagerank_big.pdf}
\caption{Scalability comparison of PageRank implementations using MapReduce,
Trilinos, and MTGL on R-MAT data sets.}
\label{f:prbig}
\end{figure}

@@

\begin{figure}[h!]
\includegraphics[width=\textwidth]{fig_cc_big.pdf}
\caption{Scalability comparison of Connected Components implementations using MapReduce,
Trilinos, and MTGL on R-MAT data sets.}
\label{f:ccbig}
\end{figure}

>>>>

In this section, we present performance results for the MapReduce
graph algorithms of the preceeding section, implemented as small C++
programs calling our MR-MPI library.
The benchmarks were run on a medium-sized Linux cluster made up of 2 GHz
dual-core AMD Opteron processors with a Myrinet network.  Most importantly
for MR-MPI, each node of the cluster has one local disk, which is used
for out-of-core operations in MR-MPI.  To avoid contention for disk I/O, we 
ran all experiments with one MPI process per node.  For comparisons with
other implementations, we used either the same cluster or, where noted,
Sandia's Cray XMT, a multi-threaded parallel computer with 500 {MHz} processors
and a 3D-Torus network.

@@

\begin{table}
\begin{center}
\begin{tabular}{|l|c|c|c|c|c|c|c|}
\hline
Data & \# of    & \# of & Maximum \\
Set  & vertices & edges & vertex degree\\
\hline
RMAT-20 (small)   &$2^{20} \approx 1M$ & $2^{23} \approx 8M$ &  $\approx 24K$ \\
RMAT-24 (medium)  &$2^{24} \approx 17M$ & $2^{27} \approx 134M$ &  $\approx 147K$ \\
RMAT-28 (large)   &$2^{28} \approx 268M$ & $2^{31} \approx 2B$& $\approx 880K$ \\
%RMAT-32 (x-large) &$2^{32} \approx 4B$ & $2^{35} \approx 34B$ &   \\
\hline
\end{tabular}
\caption{Characteristics of R-MAT input data for graph algorithm
experiments.}
\label{t:rmats}
\end{center}
\end{table}

The resulting timings give a sense of the inherent scalability of the
MapReduce algorithms as graph size grows on a fixed number of
processors, and of the parallel scalability for computing on a graph
of fixed size on a growing number of processors.  Where available, we
compare the MapReduce algorithm with other parallel implementations.
We compute parallel efficiency on $p$ processors
as $({time}_{m} \times m) / ({time}_p \times p)$
where $m$ is the smallest number of processors on which the experiment
was run, and ${time}_m$ is the execution time required on $m$ processors.

@@

\begin{figure}[htb]
\includegraphics[width=\textwidth]{fig_rmat.pdf}
\caption{Performance of R-MAT generation algorithm (Figure~\ref{fig:rmat2}).}
\label{f:rmat}
\end{figure}

@@

\begin{figure}[htb]
\includegraphics[width=\textwidth]{fig_pagerank.pdf}
\caption{Comparison of PageRank implementations using MapReduce,
Trilinos, and MTGL on R-MAT data sets.}
\label{f:pr}
\end{figure}

\subsection{Triangle finding results}

In Figure~\ref{f:tri}, we show the performance of the triangle finding 
algorithm (Figure~\ref{fig:tri}).  Execution times for this algorithm were
too large to allow the problem to be run with RMAT-28 on our cluster.
Parallel efficiencies for RMAT-20 ranged from 80\% to 140\%; for RMAT-24, they
ranged from 50\% to 120\%.

KDD NOTE:  Need to add XMT results; triangle enumeration is not implemented in PBGL.

\begin{figure}[htb]
\includegraphics[width=\textwidth]{fig_tri.pdf}
\caption{Performance of the MR-MPI triangle-finding algorithm (Figure~\ref{fig:tri}).}
\label{f:tri}
\end{figure}

\subsection{Connected Components}

We compare our MR-MPI algorithm (Figure~\ref{fig:cc}) 
with a hybrid ``Giant Connected Component''
implementation using Trilinos and MR-MPI.  Power law graphs often have one
or two very large components and many very small components.  
The hybrid algorithm exploits this feature of the data by using inexpensive
breadth-first
search from the vertex with highest degree to identify the largest components,
followed by a more expensive algorithm to identify the small components in
the remainder of the graph.  In our hybrid implementation, we use
Trilinos' matrix-vector multiplication to do the breadth-first search, 
finding components that include (in total) 70\% or more of the vertices.
We then apply our MR-MPI algorithm to the remaining graph to identify
the small components.  This hybrid approach is extremely effective,
reducing the execution time to identify all components of RMAT-20 and 
RMAT-24 by 80-99\%, as shown in Figure~\ref{f:cc}.  The benefit of MR-MPI
is seen, however, for RMAT-28, where the problem is too large to fit into
memory for our hybrid algorithm.

@@

\begin{figure}[htb]
\includegraphics[width=\textwidth]{fig_cc.pdf}
\caption{Performance of the MR-MPI connected components algorithm (Figure~\ref{fig:cc}) compared with a hybrid ``Giant Connected Component'' algorithm based
on Trilinos.}
\label{f:cc}
\end{figure}

@@

The execution times for the maximally independent set algorithm
(Figure~\ref{fig:luby}) are shown in Figure~\ref{f:luby}.  Like the R-MAT
generation results, superlinear speed-up of the algorithm occurs for RMAT-20
and RMAT-24, as more of the graph fits into processor memory and less file
I/O is needed.  For RMAT-28, the algorithm requires significant out-of-core
operations. In this case, parallel efficiency is nearly perfect going from 
8 to 32 processors; some loss of efficiency is seen at 64 processors,
where the parallel efficiency compared to the 8-processor case is KDD\%.

KDD NOTE:  I think I had a bad node for the 64-processor run; I am 
rerunning it now and soon will have revised RMAT-28 results on 64 processors.

@@

\begin{figure}[htb]
\includegraphics[width=\textwidth]{fig_luby.pdf}
\caption{Performance of maximally independent set algorithm (Figure~\ref{fig:luby}).}
\label{f:luby}
\end{figure}

@@

\begin{figure}[htb]
\includegraphics[width=\textwidth]{fig_sssp.pdf}
\caption{Execution times for SSSP using MR-MPI with R-MAT matrices.  
Both the original algorithm (Figure~\ref{fig:sssp}) and the enhanced 
algorithm (Figure~\ref{fig:sssp2}) are included.}
\label{f:sssp}
\end{figure}

@@

\begin{figure}[htb]
\includegraphics[width=\textwidth]{fig_ssspA.pdf}
\caption{SSSP using MR-MPI for WebGraphA with
13.2M vertices and 31.9M edges.  Runtimes using Hadoop and PBGL
are also shown.}
\label{f:ssspA}
\end{figure}

\begin{table}
\begin{center}
\begin{tabular}{|l|c|r|}
\hline
Implementation & Number of Processes & SSSP Execution Time \\
\hline
Hadoop & 48  & 38,925 secs.\\
MR-MPI original (Figure~\ref{fig:sssp}) & 48 &  13,505 secs.\\
MR-MPI enhanced (Figure~\ref{fig:sssp2}) & 48 &  8,031 secs.\\
XMT/C  & 32 &  37 secs.\\
%Out-of-core MR-MPI & 96 &   8,358 secs.\\
%Out-of-core MR-MPI & 64 &  12,882 secs.\\
%Out-of-core MR-MPI & 100 &  6,280 secs.\\
\hline
\end{tabular}
\caption{Execution times for SSSP with {WebGraphB}.}
\label{t:ssspB}
\end{center}
\end{table}

\subsection{Scalability to large numbers of processors}
Finally, we demonstrate the scalability of MR-MPI to large numbers of 
processors.  
The MR-MPI implementation was run on Sandia's Redstorm and Thunderbird
parallel computers.  Redstorm is a massively parallel computer with 2+ 
{GHz} dual/quad-core AMD Opteron processors and 9.6 GB/s link bandwidth.
Thunderbird is a cluster with 3.6 {GHz} dual-core Intel EM64T processors
and an Infiniband network.
Because these systems do not have local disks for each processor, we selected 
a data set and page sizes that fit in memory, so out-of-core operations
were not used.  For these experiments, we used an R-MAT data set with 
with $2^{25}$ vertices and $2^{28}$ edges, with parameters given in
Table~\ref{t:rmat}.  We ran both PageRank and Connected Components
algorithms.

\begin{table}
\begin{center}
\begin{tabular}{|l|c|c|c|c|c|c|c|}
\hline
Data & R-MAT  & R-MAT  & R-MAT  & R-MAT  & \# of    & \# of & Maximum \\
Set  & a      & b      & c      & d      & vertices & edges & vertex degree\\
\hline
nice  & 0.45 & 0.15 & 0.15 & 0.25 & $2^{25}$ & $2^{28}$ & 1108 \\
nasty & 0.57 & 0.19 & 0.19 & 0.05 & $2^{25}$ & $2^{28}$ & 230,207\\
\hline
\end{tabular}
\caption{Characteristics of R-MAT input data for PageRank and Connected
Components scalability experiments.}
\label{t:rmat}
\end{center}
\end{table}

@@

\begin{figure}[htb]
\includegraphics[width=\textwidth]{fig_pagerank_big.pdf}
\caption{Scalability comparison of PageRank implementations using MapReduce,
Trilinos, and MTGL on R-MAT data sets.}
\label{f:prbig}
\end{figure}

@@

\begin{figure}[htb]
\includegraphics[width=\textwidth]{fig_cc_big.pdf}
\caption{Scalability comparison of Connected Components implementations using MapReduce,
Trilinos, and MTGL on R-MAT data sets.}
\label{f:ccbig}
\end{figure}
